---
title: "Ecommerce Customers Churn Analysis"
author: "Ravina Ingole, Abhishek Kukreja and Akash Dhage "
date: "`r format(Sys.time(),'%B %d, %Y')`"
output: 
      html_document:
        toc: true
        toc_float:
          collapse: false
          smooth_scroll: true
---

<style>
#TOC {
  color: blue;
  font-family: Calibri;
  font-size: 16px; 
  border-color: #708090;
}
#header {
  color: black;
  background-color: #F5F5F5;
  opacity: 0.6;
  font-family: Calibri;
  font-size: 20px;
}
body {
  color: #708090;
  font-family: Calibri;
  background-color: #F5F5F5;
}
pre {
  color: #708090;
  background-color: #F8F8FF;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data, include=FALSE}
#library(tidyverse)
library(ggthemes)
library(wesanderson)
library(ggplot2)
library(datasets)
library(caTools)
library(dplyr)
library(magrittr)
library(party)

# reading contents of csv file
hd <- read.csv("C:/Users/Abhishek/Desktop/Abhishek/Sem 2/DA/Project/Ecommerce/Ecommerce Dataset.csv")
```

# Summary of `E-commerce dataset` and cleaning.

```{r Dataset Summary}
print(head(hd))
print(str(hd))
print(summary(hd))
```
### Finding number of null values.

```{r Finding number of null values}
print(colSums(is.na(hd)))

```

### Checking Outliers

```{r Checking outliers using boxplot}
boxplot(hd$Tenure, hd$WarehouseToHome, hd$DaySinceLastOrder, hd$NumberOfAddress,
main = "Multiple boxplots for comparision",
at = c(1,2,4,5),
names = c("Tenure", "WarehouseToHome", "DaySinceLastOrder", "NumberOfAddress"),
las = 2,
col = c("orange","red","Black", "Grey"),
border = "brown",
notch = TRUE
)
```


```{r Checking outliers using boxplot for different variables}
boxplot(hd$OrderAmountHikeFromlastYear, hd$CouponUsed, hd$OrderCount,
main = "Multiple boxplots for comparision",
at = c(1,2,4),
names = c("OrderAmountHikeFromlastYear", "CouponUsed", "OrderCount"),
las = 2,
col = c("orange","red","light Blue"),
border = "brown",
notch = TRUE
)
```
```{r Checking outliers using boxplot for CashbackAmount}
boxplot(hd$CashbackAmount,data=hd, main="Churn Vs Cash Back Amount",
   xlab="Churn", ylab="Cash Back Amount")
```


```{r Checking outliers using boxplot for OrderAmountHikeFromlastYear}
boxplot(hd$OrderAmountHikeFromlastYear,data=hd, main="Order Amount Hike From last Year",
   xlab="Churn", ylab="Order Amount Hike From last Year")
```


### Plotting Histogram

```{r Histogram for Tenure}
ggplot(hd, aes(x=Tenure))+
  geom_histogram(color="darkblue", fill="lightblue")
```

```{r Histogram for Warehouse To Home}
ggplot(hd, aes(x=WarehouseToHome))+
  geom_histogram(color="darkblue", fill="lightblue")
```

```{r Histogram for Hour Spend On App}
ggplot(hd, aes(x=HourSpendOnApp))+
  geom_histogram(color="darkblue", fill="lightblue")
```

```{r Histogram for Coupon Used}
ggplot(hd, aes(x=CouponUsed))+
  geom_histogram(color="darkblue", fill="lightblue")
```


```{r Histogram for Order Count}
ggplot(hd, aes(x=OrderCount))+
  geom_histogram(color="darkblue", fill="lightblue")
```

```{r Histogram for Day Since Last Order}
ggplot(hd, aes(x=DaySinceLastOrder))+
  geom_histogram(color="darkblue", fill="lightblue")
```
```{r Histogram for Order Amount Hike From last Year}
hist <-ggplot(hd, aes(x=OrderAmountHikeFromlastYear))
  hist + geom_histogram(color="darkblue", fill="lightblue")
```


### Plotting Box Plot

```{r Box plot of churn}
counts <- table(hd$Churn)
barplot(counts, main="Churn Distribution",
   xlab="Churn")
```
```{r Box plot of SatisfactionScore}
tbl <- with(hd, table(Churn, SatisfactionScore))
barplot(tbl, beside = TRUE, legend = TRUE,main="Satisfaction Score",
   xlab="Satisfaction")
```
```{r Box plot of Tenure and Churn}
tbl <- with(hd, table(Churn, Tenure))
barplot(tbl, beside = TRUE, legend = TRUE,main="Impact of Tenure on Churn",
   xlab="Tenure")
```

```{r Box plot of HourSpendOnApp and Churn}
tbl <- with(hd, table(Churn, HourSpendOnApp))
barplot(tbl, beside = TRUE, legend = TRUE,main="Hours Spent on App/Website Distribution",
   xlab="HourSpendOnApp")
```

```{r Box plot of Complain and Churn}
tbl <- with(hd, table(Churn, Complain))
barplot(tbl, beside = TRUE, legend = TRUE,main="Impact of No. of Complains on Churn",
   xlab="Complain")
```

```{r Box plot of PreferredPaymentMode and Churn}
tbl <- with(hd, table(Churn, PreferredPaymentMode))
barplot(tbl, beside = TRUE, legend = TRUE,main="Payment Mode vs Churn",
   xlab="PreferredPaymentMode")
```

```{r Box plot of Gender and Churn}
tbl <- with(hd, table(Churn, Gender))
barplot(tbl, beside = TRUE, legend = TRUE,main="Gender-wise Churn Distribution",
   xlab="Gender")
```

```{r Box plot of CityTier and Churn}
tbl <- with(hd, table(Churn, CityTier))
barplot(tbl, beside = TRUE, legend = TRUE,main="City Tier Churn Distribution",
   xlab="CityTier")
```

### Plotting scatter plot

```{r Scatter plot of Tenure and CashbackAmount}
plot(hd$Tenure, hd$CashbackAmount, col='grey',main="Tenure vs Cashback Amount", pch=10)

```

```{r Scatter plot of Order Count vs Coupon Used}
plot(hd$OrderCount, hd$CouponUsed, col='grey',main="Order Count vs Coupon Used", pch=10)

```

```{r Scatter plot of Coupon Used vs Day Since Last Order}
plot(hd$DaySinceLastOrder, hd$CouponUsed, col='grey',main="Coupon Used vs Day Since Last Order", pch=10)

```

```{r Scatter plot of Order Count vs Day since Last Order}
plot(hd$DaySinceLastOrder, hd$OrderCount, col='grey',main="Order Count vs Day since Last Order", pch=10)

```

# Handling Null Values

```{r Handling null values with median}
medTenure <- median(hd$Tenure,na.rm = TRUE)
medWarehouseToHome <- median(hd$WarehouseToHome,na.rm = TRUE)
medHourSpendOnApp <- median(hd$HourSpendOnApp,na.rm = TRUE)
medCouponUsed <- median(hd$CouponUsed,na.rm = TRUE)
medOrderAmountHikeFromlastYear <- median(hd$OrderAmountHikeFromlastYear,na.rm = TRUE)
medDaySinceLastOrder <- median(hd$DaySinceLastOrder,na.rm = TRUE)
medOrderCount <- median(hd$OrderCount,na.rm = TRUE)

hd[is.na(hd$Tenure),'Tenure'] <- medTenure
hd[is.na(hd$WarehouseToHome),'WarehouseToHome'] <- medWarehouseToHome
hd[is.na(hd$HourSpendOnApp),'HourSpendOnApp'] <- medHourSpendOnApp
hd[is.na(hd$CouponUsed),'CouponUsed'] <- medCouponUsed
hd[is.na(hd$OrderAmountHikeFromlastYear),'OrderAmountHikeFromlastYear'] <- medOrderAmountHikeFromlastYear
hd[is.na(hd$DaySinceLastOrder),'DaySinceLastOrder'] <- medDaySinceLastOrder
hd[is.na(hd$OrderCount),'OrderCount'] <- medOrderCount
print(colSums(is.na(hd)))
```


# Decisssion Tree

### Spliting data into test and train data 

```{r Spliting data into test and train data}
library(rsample)  # testing/training splits 
library(recipes)     # preprocessing
library(yardstick)   #  ROC/AUC
create_train_test <- function(data, size = 0.8, train = TRUE) {
    n_row <- nrow(data)
    total_row <- size * n_row
    train_sample <- 1:total_row
    if (train == TRUE) {
        return (data[train_sample, ])
      } else {
        return (data[-train_sample, ])
     }
}
data_train <- create_train_test(hd, 0.8, train = TRUE)
data_test <- create_train_test(hd, 0.8, train = FALSE)
print(dim(data_train))
print(dim(data_test))
```

### Desicion tree  

```{r Desicion Tree}
library(rpart)
library(rpart.plot)
fit <- rpart(Churn~., data = data_train, method = 'class')
rpart.plot(fit,extra= 106)

#default.ct <- rpart(Personal.Loan ~ ., data = train.df, method = "class")
# plot tree
prp(fit, type = 1, extra = 1, under = TRUE, split.font = 1, varlen = -10)


```

### Making a Prediction  

```{r Making a Prediction}
predict_unseen <-predict(fit, data_test, type = 'class')
```

### Testing the churn customers  

```{r Testing the churn customers}
table_mat_test <- table(data_test$Churn, predict_unseen)
table_mat_test
```

### Confusion Matrix  

```{r Confusion Matrix}
accuracy_Test <- sum(diag(table_mat_test)) / sum(table_mat_test)
print(paste('Accuracy for test for Decission tree', accuracy_Test))
```
### Results for Decision Tree
Accuracy: 89%

Precision: 78%

Recall: 53%

F1 Score: 63%


### Complexity Parameter

```{r pa_cp}
cv.ct <- rpart(Churn ~ ., data = data_train, method = "class", 
    cp = 0.00001, minsplit = 5, xval = 5)
printcp(cv.ct)

pruned.ct <- prune(cv.ct,
cp = cv.ct$cptable[which.min(cv.ct$cptable[,"xerror"]),"CP"])
length(pruned.ct$frame$var[pruned.ct$frame$var == "<leaf>"])
prp(pruned.ct, type = 1, extra = 1, split.font = 1, varlen = -10)

predict_unseen <-predict(pruned.ct, data_test, type = 'class')

table_mat_test <- table(data_test$Churn, predict_unseen)
table_mat_test

accuracy_Test <- sum(diag(table_mat_test)) / sum(table_mat_test)
print(paste('Test Accuracy for Pruned Decission tree', accuracy_Test))
```
### Results for Pruned Decision Tree
Accuracy: 95%

Precision: 90%

Recall: 72%

F1 Score: 80%


# Random Forest  

### Random Forest training model

```{r Random Forest training model}
library(randomForest)
set.seed(200)
data_train_random <- create_train_test(hd, 0.8, train = TRUE)
data_test_random <- create_train_test(hd, 0.8, train = FALSE)
hd_rf = randomForest(factor(Churn)~., data=data_train_random, ntree=500, proximity=T)
table(predict(hd_rf), data_train_random$Churn)
hd_rf
plot(hd_rf)
```
### Finding Feature Importance

```{r Finding Feature Importance}
importance(hd_rf)

```
### Making a Prediction  

```{r Making a Prediction rf}
predict_unseen_rf <-predict(hd_rf, data_test_random, type = 'class')
```

### Testing the churn customers  

```{r Testing the churn customers rf}
table_mat_rf <- table(data_test_random$Churn, predict_unseen_rf)
table_mat_rf
```


### Confusion Matrix  

```{r Confusion Matrix rf}
accuracy_Test_rf <- sum(diag(table_mat_rf)) / sum(table_mat_rf)
print(paste('Accuracy for test for Decission tree', accuracy_Test_rf))

```
### Results for Random Tree
Accuracy: 96%

Precision: 100%

Recall: 71%

F1 Score: 83%

# Conclusion

Basis F1 score, Random Forest has better prediction power compared to Decision Tree and Pruned Decision Tree.






